---
title: "Machine Learning Exercise Regime"
author: "Simeon"
date: "Saturday, August 22, 2015"
output: html_document
---

## Executive Summary

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively.  One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, the goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

# Experimental Setup

Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes. Participants were supervised by an experienced weight lifter to make sure the execution complied to the manner they were supposed to simulate. The exercises were performed by six male participants aged between 20-28 years, with little weight lifting experience. We made sure that all participants could easily simulate the mistakes in a safe and controlled manner by using a relatively light dumbbell (1.25kg).

We will use Decision Trees to conduct the modelling of the test results.


# Loading and Cleaning the Training and Test Data

```{r, echo=FALSE}

setwd("C:/Users/Simeon/Desktop/R Language")

# Loading Training set

if(!file.exists("./machinelearning")){dir.create("./machinelearning")}
filename <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
download.file(filename, destfile = "./machinelearning/Trainingset.csv")
Training <- read.csv("C:/Users/Simeon/Desktop/R Language/machinelearning/Trainingset.csv", na.strings=c("NA","#DIV/0!",""))


setwd("C:/Users/Simeon/Desktop/R Language/machinelearning")

# Loading Test set

if(!file.exists("./machinelearning")){dir.create("./machinelearning")}
filename <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(filename, destfile = "./machinelearning/Testset.csv")
Test <- read.csv("C:/Users/Simeon/Desktop/R Language/machinelearning/Testset.csv", na.strings=c("NA","#DIV/0!",""))

```

# Partitioning the Training Set and removing Excess NA values

```{r, echo=TRUE}
library(caret)
library(randomForest)

PartitionedTrain <- createDataPartition(y=Training$classe, p=0.6, list=FALSE)
NewTraining <- Training[PartitionedTrain, ]
NewTesting <- Training[-PartitionedTrain, ]

DuplicateTraining <- NewTraining

for(i in 1:length(NewTraining)) {
        if( sum( is.na( NewTraining[, i] ) ) /nrow(NewTraining) >= .6 ) {
        for(j in 1:length(DuplicateTraining)) {
            if( length( grep(names(NewTraining[i]), names(DuplicateTraining)[j]) ) ==1)  { 
                DuplicateTraining <- DuplicateTraining[ , -j]
            }   
        } 
    }
}

NZV <- nearZeroVar(NewTraining, saveMetrics=TRUE)
if (any(NZV$nzv)) nzv else message("Variables have no near zero variance")

```

# Building a Decision Tree

```{r, echo=TRUE}

library(rpart) 
RPartitionedFit <- rpart(classe ~ ., data=NewTraining, method="class")
library(rattle)
fancyRpartPlot(RPartitionedFit)
```
